---
title: "Export proportion of geocoded data per commune"
output: html_document
---

NOTE: this file is deprecated, refer to `codes/99_posthoc_check.R` instead

```{r}
library(tidyverse)
library(qs2)
```

## Summary

Create a dataset summarizing proportion of geocoded data, grouping by: date hospitalized + commune (by GADM convention)

## Data

### Load data

Load minimal data

```{r}
dat_2000_2016 <- read_csv("./data/incidence/incidence_2000_2016_minimal.csv")
dat_2017_2025 <- read_csv("./data/incidence/incidence_2017_2025_minimal.csv")

full_incidence <- bind_rows(
  dat_2000_2016,
  dat_2017_2025
)
full_incidence
```

Load cleaned geocoded addresses

```{r}
geocoded_addr_clean <- qs_read("./data/cached/geocoded_addr_clean.qs")
geocoded_addr_clean
```

Load geo data

```{r}
hcmc_shape <- read_rds("./data/gadm/hcmc_shapefiles.rds")
commune_gadm <- hcmc_shape$commune_322 
commune_gadm
```

### Clean communes

Compute the proportion of geocoded data

```{r}
full_incidence_clean <- full_incidence %>% 
  left_join(
    geocoded_addr_clean %>% 
      select(raw_addr,vietmap_district, vietmap_commune, long, lat)
  ) %>% 
  filter(
    ng_vaovien >= as.Date("2002-01-01")
  ) %>% 
  mutate(
    # Format px, qh such that it matches that of GADM
    px = str_remove(px, "(Phường |Xã |Thị trấn |Thị Trấn |THI TRAN)"),
    px = case_when(
      # handle px starts with 0
      str_detect(px, "^0\\d$") ~ str_replace(px, "0", "Phường "),
      str_detect(px, "^\\d+$") ~ paste0("Phường ", px),
      .default = px
    ),
    px = stri_trans_general(px, "Latin-ASCII"),
    px = str_to_title(px),
    qh = str_remove(qh, "(Quận |Huyện )"),
    qh = case_when(
      # handle px starts with 0
      str_detect(qh, "^0\\d$") ~ str_replace(qh, "0", "Quận "),
      str_detect(qh, "^\\d+$") ~ paste0("Quận ", qh),
      .default = qh
    ),
    qh = stri_trans_general(qh, "Latin-ASCII"),
    qh = str_to_title(qh),
    # Do the exact same thing w/ Vietmap district and commune
    vietmap_district = str_remove(vietmap_district, "(Quận |Huyện )"),
    vietmap_district = if_else(
      str_detect(vietmap_district, "^\\d+$"), 
      paste0("Quận ", vietmap_district), vietmap_district
    ),
    vietmap_district = stri_trans_general(vietmap_district, "Latin-ASCII"),
    vietmap_district = str_to_title(vietmap_district),
    vietmap_commune = str_remove(vietmap_commune, "(Phường |Xã |Thị Trấn )"),
    vietmap_commune = if_else(
      str_detect(vietmap_commune, "^\\d+$"), 
      paste0("Phường ", vietmap_commune), vietmap_commune
    ),
    vietmap_commune = stri_trans_general(vietmap_commune, "Latin-ASCII"),
    vietmap_commune = str_to_title(vietmap_commune),
  ) 
```

```{r}
# update qh Khong Ro and px Khong Ro with VietMap output if available
full_incidence_clean <- full_incidence_clean %>% 
  mutate(
    qh = if_else((is.na(qh) | qh == "Khong Ro" | qh == ""), vietmap_district, qh),
    px = if_else((is.na(px) | px == "Khong Ro" | px == ""), vietmap_commune, px)
  ) 
```

### Get corresponding GADM name

Map qh, px to corresponding GADM varname_3 for consistency

```{r}
commune_gadm <- commune_gadm %>% 
  select(name_2, name_3, varname_3, geometry) %>% 
    mutate(
      # generate key for mapping
      name_2_key = stri_trans_general(name_2, "Latin-ASCII"),
      name_2_key = str_to_title(name_2_key),
      # generate key for mapping
      name_3_key = stri_trans_general(name_3, "Latin-ASCII"),
      name_3_key = str_to_title(name_3_key)
    ) 
```

First, try mapping by name

```{r}
full_incidence_gadm <- full_incidence_clean %>% 
  left_join(
    commune_gadm %>% 
      as_tibble() %>% 
      select(-geometry),
    by = join_by(
      px == name_3_key, qh == name_2_key
    )
  )

# get the incidence that have been mapped to GADM commune name here
full_incidence_gadm_1 <- full_incidence_gadm %>% 
  filter(!is.na(name_3))
```

For communes that failed to map using name, try using geocoded data for mapping (if available)

```{r}
full_incidence_gadm_2 <- full_incidence_gadm %>% 
  filter(is.na(name_3)) %>% 
  select(-name_3, -name_2, -varname_3) %>% 
  filter(!is.na(long)) %>% 
  mutate(
    long_pt = long,
    lat_pt = lat
  ) %>% 
  st_as_sf(coords = c("long_pt", "lat_pt"), crs = crs(commune_gadm)) %>% 
  st_join( commune_gadm %>% 
             select(name_2, name_3, varname_3), 
           join = st_within) %>% 
  as_tibble() %>% 
  select(-geometry) %>% 
  filter(
    !is.na(name_3)
  )

full_incidence_gadm_2
```

**Note** 3,704 incidence cannot be geocoded and does not have data up to district level

```{r}
full_incidence_gadm %>% 
  filter(is.na(name_3), is.na(long))
```

### Compute geocoded prop

Summarize based on name_3 in GADM

```{r}
geocoded_prop <- bind_rows(
  full_incidence_gadm_1,
  full_incidence_gadm_2
) %>% 
  group_by(
    ng_vaovien, name_2, name_3, varname_3
  ) %>% 
  summarize(
    total_inc = n(),
    na_address = sum(is.na(diachi)),
    geocoded_inc = sum(!is.na(long)),
    geocoded_prop = geocoded_inc/total_inc
  ) %>% 
  ungroup()
```

```{r}
# geocoded_prop %>%
#   rename(
#     date_hosp = ng_vaovien
#   ) %>%
#   write_rds("./data/export/geocoded_prop.rds")

read_rds("./data/export/geocoded_prop.rds")
```

```{r}
full_incidence_gadm %>% 
  filter(is.na(name_3), is.na(long)) %>% 
  rename(
    raw_address = diachi,
    raw_district = qh, 
    raw_commune = px,
    hosp_date = ng_vaovien
  ) %>% 
  select(raw_address, raw_commune, raw_district, hosp_date) %>% 
  write_rds("./out/data/failed_gadm_map.rds")
```
