---
title: "HCMC, Ba Ria - Vung Tau, Binh Duong shapefiles"
output: html_document
---

```{r}
library(tidyverse)
library(sf)
library(terra)
library(janitor)
library(stringi)
library(readxl)
```

Check the shapefiles from HCMC, Ba Ria - Vung Tau

```{r}
# Helper functions

# generate joining key for px, qh name
make_key <- function(x) {
  x %>% 
    str_trim() %>% 
    stri_trans_general("Latin-ASCII") %>% 
    str_to_lower() %>% 
    str_replace_all("\\s+", "_")
}

# handle sap_nhap column to get a map from old commune to new commune
parse_merge_fr <- function(sap_nhap_str, ten_xa_moi) {
  entries <- str_split(sap_nhap_str, ",\\s*(?![^(]*\\))")[[1]]
  
  map_dfr(entries, function(entry) {
    entry <- str_trim(entry)
    
    # extract all tokens inside parentheses, split by comma
    paren_content <- str_extract(entry, "(?<=\\()([^)]+)(?=\\))")
    tokens <- if (!is.na(paren_content)) str_trim(str_split(paren_content, ",")[[1]]) else character(0)
    
    # classify each token
    is_partial    <- any(str_detect(tokens, "(?i)^1 ph[aầ]n$"))
    district_token <- tokens %>% 
          keep(~ str_detect(.x, "(?i)^(qu[aậ]n|huy[eệ]n|th[iị] x[aã]|th[àa]nh ph[oố])"))  %>% 
  str_remove("(?i)^(qu[aậ]n|huy[eệ]n|th[iị] x[aã]|th[àa]nh ph[oố])\\s+")
    
    tibble(
      ten_xa_cu     = str_remove_all(entry, "\\s*\\([^)]*\\)") %>%  str_trim(),
      ten_huyen_cu   = if (length(district_token) > 0) district_token[[1]] else NA_character_,
      ten_xa_moi    = ten_xa_moi,
      loai_sap_nhap = if_else(is_partial, "1 phan", "toan phan")
    )
  }) 
}
```

## Shape files

```{r}
hcmc_shapefiles <- read_rds("../data/gadm/hcmc_shapefiles.rds")
```

### GISvn

Map after the 2021 reform

```{r}
commune_gis <- read_sf("../data/shapefiles/before_reforms/gisvn_vnm_shp/Việt Nam (phường xã) - 63.shp", options = "ENCODING=UTF-8") %>% 
  clean_names() 

commune_gis <- commune_gis %>% 
  filter(
    ten_tinh == "TP. Hồ Chí Minh" |
    ten_tinh == "Bà Rịa - Vũng Tàu" |
    ten_tinh == "Bình Dương"
  ) %>% 
  mutate(
    ten_xa = str_remove(ten_xa, "^0+(?=[1-9])")
  )
```

```{r}
commune_gis %>% 
  group_by(ten_tinh) %>% 
  summarize(
    no_commune = n()
  )
```

### Get the standard name

Create a data.frame with the standard commune names and a mapping to new communes

```{r}
commune_names_gis <- commune_gis %>% 
  as_tibble() %>% 
  select(-cap, -stt, -geometry) %>% 
  arrange(ten_tinh)
```

```{r}
commune_merged <- read_sf("../data/shapefiles/after_2025_reforms/Việt Nam (phường xã) - 34/", options = "ENCODING=UTF-8") %>% 
  clean_names() %>% 
  filter(
    ten_tinh == "TP. Hồ Chí Minh" 
  ) 
  
commune_names_merged <- commune_merged %>% 
  as_tibble() %>% 
  select(ma_xa, ten_xa, sap_nhap, loai, ma_tinh)
```

Generate a data.frame mapping from old commune to new commune

```{r}
old_to_new_map <- commune_names_merged  %>% 
  select(ten_xa, sap_nhap)  %>% 
  pmap_dfr(~ parse_merge_fr(..2, ..1)) %>% 
  ungroup() %>% 
  mutate(
    ten_xa_cu = str_remove(ten_xa_cu, "Phường\\s+")
  )

# save a csv to adjust manually
# write_csv(old_to_new_map, "../data/shapefiles/for_HCDC/px_cu_moi_map.csv")

old_to_new_map <- read_csv("../data/shapefiles/for_HCDC/px_cu_moi_map.csv") %>% 
  mutate(
    key_xa_cu = make_key(ten_xa_cu),
    key_huyen_cu = make_key(ten_huyen_cu)
  )
```

Finally, merge to get the full details of old and new communes

```{r}
# ---- add details of old communes
old_new_map_full <- bind_rows(
  old_to_new_map %>%  
    filter(!is.na(ten_huyen_cu)) %>%  
    left_join(commune_names_gis %>% 
                select(-loai) %>%
                mutate(
                  ten_huyen = make_key(ten_huyen),
                  ten_xa = make_key(ten_xa)
                ) %>% 
                rename(
                  ma_xa_cu = ma_xa,
                  ma_huyen_cu = ma_huyen,
                  ma_tinh_cu = ma_tinh,
                  ten_tinh_cu = ten_tinh
                ), 
              by = join_by(
                key_huyen_cu == ten_huyen,
                key_xa_cu == ten_xa)
              ),
  old_to_new_map %>%  
    filter(is.na(ten_huyen_cu)) %>%  
    select(-ten_huyen_cu) %>% 
    left_join(commune_names_gis %>% 
                select(-loai) %>% 
                mutate(
                  ten_xa = make_key(ten_xa)
                ) %>%
                rename(
                  ma_xa_cu = ma_xa,
                  ten_huyen_cu = ten_huyen,
                  ma_huyen_cu = ma_huyen,
                  ten_tinh_cu = ten_tinh,
                  ma_tinh_cu = ma_tinh
                ), 
              by = join_by(key_xa_cu == ten_xa)
              )
)

# ---- add details of new communes
old_new_map_full <- old_new_map_full %>% 
  left_join(
    commune_names_merged %>% 
      select(ma_xa, ten_xa) %>% 
      rename(
        ten_xa_moi = ten_xa,
        ma_xa_moi = ma_xa
      )
  ) %>% 
  select(
    ten_xa_cu, ma_xa_cu, ten_huyen_cu, ma_huyen_cu, ten_tinh_cu, ma_tinh_cu,
    ten_xa_moi, ma_xa_moi, loai_sap_nhap
  )
```

Write necessary files

```{r}
write_csv(old_new_map_full, "../data/shapefiles/for_HCDC/px_cu_moi.csv")
shp_files <- list(
  before_merge = commune_gis,
  after_merge = commune_2025
)
write_rds(shp_files, "../data/shapefiles/for_HCDC/new_hcm_shapefiles.rds")
shp_files$before_merge
shp_files$after_merge
```

Save communes that exists in shapefile but **NOT** in old to new mapping

This mean that these communes were changed during the 2023-2025 period

-   Example: Q8 as detailed [here](https://ghtk.vn/blog/phuong-1-quan-8-sap-nhap/)

```{r}
commune_gis %>% 
  as_tibble() %>% 
  select(ma_xa, ten_xa, ten_huyen, ma_huyen, ten_tinh, ma_tinh) %>% 
  left_join(
    old_new_map_full %>% 
      select(ma_xa_cu, ten_xa_cu),
    by = join_by(ma_xa == ma_xa_cu)
  ) %>% 
  filter(is.na(ten_xa_cu)) %>% 
  select(-ten_xa_cu) %>% 
  arrange(ten_tinh, ten_huyen, ten_xa) %>% 
  write_csv("../data/shapefiles/for_HCDC/thay_doi_trong_2023_2025.csv")
```

## Incidence data

```{r}
old_area_raw_incidence_2017_2025 <- bind_rows(
  map(excel_sheets("../data/incidence/incidence_2017_2025.xlsx"),
      ~ read_excel("../data/incidence/incidence_2017_2025.xlsx", sheet = .x)),
  map(excel_sheets("../data/incidence/incidence_bd_vt_2017_2025.xlsx"),
      ~ read_excel("../data/incidence/incidence_bd_vt_2017_2025.xlsx", sheet = .x))
)
```

## Helper functions

```{r}
source("../R/changes_in_admin_lvl3.R")
```

### Data to gis sf 

```{r}
generate_key_dat_to_2023 <- function(data, gis){
  newdat <- data %>% 
    mutate(
      city_key     = make_key(thanh_pho_cu),
      city_key     = recode(city_key, "tp._h.c.m"="tp._ho_chi_minh"),
      city_key     = str_remove(city_key, "^tinh_"),
      district_key = make_key(quan_huyen_cu),
      district_key = str_remove(district_key, "^(quan_|huyen_|thanh_pho_)"),
      commune_key  = make_key(phuong_xa_cu),
      commune_key = str_remove(commune_key, "^(phuong_|xa_|thi_tran_)")
    ) %>% 
    group_by(commune_key, district_key, city_key) %>% 
    # ------ handle changes in district/commune between data & 2023 GIS sf --------
    left_join(commune_district_crosswalk_2023,
              by = c("city_key", "district_key", "commune_key"),
              relationship = "many-to-one") %>% 
    mutate(
      district_key = coalesce(shapefile_district_key, district_key),
      commune_key = coalesce(shapefile_commune_key, commune_key)
    )
  
  newgis <- gis %>% 
    mutate(
      city_key     = make_key(ten_tinh),
      district_key = make_key(ten_huyen),
      commune_key  = make_key(ten_xa)
    ) %>% 
    hcmc_shapefiles$merge_polygon(
      merge_map = merge_gis_polygons,
      colname = "commune_key"
    ) %>% 
    # special handling for tam_an
    mutate(
      city_key = if_else(commune_key == "tam_an", 
                         "ba_ria_-_vung_tau",
                         city_key),
      district_key = if_else(commune_key == "tam_an", 
                         "long_dat",
                         district_key)
    )
  
  list(
    dat = newdat,
    sf = newgis
  )
}
```

Test joining modified incidence and shapefile

```{r}
test_merge <- generate_dat_to_gissf(
  old_area_raw_incidence_2017_2025,
  shp_files$before_merge
)

full_join(
  test_merge$dat %>% 
    select(city_key, district_key, commune_key) %>% 
    unique(),
  test_merge$sf,
  by = c("city_key", "district_key", "commune_key"),
  relationship = "many-to-one"
)
```

```{r}
full_join(
  test_merge$dat,
  test_merge$sf,
  by = c("city_key", "district_key", "commune_key"),
  relationship = "many-to-one"
)
```

### Data to new commune 

```{r}
generate_key_dat_to_2025 <- function(data, 
                                     sf_dat){
  newdat <- data %>% 
    mutate(
      city_key     = make_key(thanh_pho_cu),
      city_key     = recode(city_key, "tp._h.c.m"="tp._ho_chi_minh"),
      city_key     = str_remove(city_key, "^tinh_"),
      district_key = make_key(quan_huyen_cu),
      district_key = str_remove(district_key, "^(quan_|huyen_|thanh_pho_)"),
      commune_key  = make_key(phuong_xa_cu),
      commune_key = str_remove(commune_key, "^(phuong_|xa_|thi_tran_)")
    ) %>% 
    group_by(commune_key, district_key, city_key) %>% 
    # ------ step 1: map data district/commune to 2023 GIS sf key --------
    left_join(commune_district_crosswalk_2023,
              by = c("city_key", "district_key", "commune_key"),
              relationship = "many-to-one") %>% 
    mutate(
      district_key = coalesce(shapefile_district_key, district_key),
      commune_key = coalesce(shapefile_commune_key, commune_key)
    ) %>% 
    # ------ step 2: map data from 2023 key to 2025 GIS sf key -------
    left_join(commune_district_crosswalk_2023_2025,
              by = c("city_key", "district_key", "commune_key"),
              relationship = "many-to-one") %>% 
    mutate(
      district_key = coalesce(new_district_key, district_key),
      commune_key = coalesce(new_commune_key, commune_key)
    ) 
  
  list(
    dat = newdat,
    sf = sf_dat
  )
}
```

Try mapping from commune/district in incidence data to a commune after merge

-   If all 3 keys (commune, district, city) r available –\> join by all 3 keys

-   Otherwise, fall back to joining using commune only

```{r}
test_merge <- generate_key_dat_to_2025(old_area_raw_incidence_2017_2025, 
                                       old_new_map_full)

map_full_keyed <- old_new_map_full %>%
  mutate(
    commune_key  = make_key(ten_xa_cu),
    district_key = make_key(ten_huyen_cu),
    city_key     = make_key(ten_tinh_cu)
  )

three_key <- test_merge$dat %>% 
  select(city_key, district_key, commune_key) %>% 
  unique() %>% 
  left_join(map_full_keyed %>% filter(!is.na(district_key), !is.na(city_key)),
            by = c("city_key", "district_key", "commune_key"))

bind_rows(
  three_key %>% filter(!is.na(ten_xa_cu)),  # matched
  three_key %>% filter( is.na(ten_xa_cu)) %>%  # unmatched → try fallback
    select(city_key, district_key, commune_key) %>%
    left_join(map_full_keyed %>% filter(is.na(district_key), is.na(city_key)) %>%
                select(-city_key, -district_key), by = "commune_key")
)
```
